---
title: "Colour experiment"
output: pdf_document
---
# Introduction

This analysis looks at the sign variants used in a colour naming game between signers of different sign languages meeting after 1 week of interaction and after 3 weeks of interaction.  The data was collected by Kang Suk Byun (Kang-Suk.Byun@mpi.nl).

The analysis tries to predict the relative frequency of each variant within a colour category in week 3, based on measures from week 1.

# Load libraries
```{r warning=F, message=F}

library(ggplot2)
library(lme4)
library(party)
library(Rmisc)
library(dplyr)

```

```{r echo=F}

setwd("~/Documents/MPI/KangSukColours/ColourExperiment/analysis/")

getMEText = function(r,ef, wald=NULL){
  
  AIC = r[2,]$AIC
  loglikDiff = signif(diff(r$logLik),2)
  chi = round(r$Chisq[2],2)
  df = r$`Chi Df`[2]
  p = signif(r$`Pr(>Chisq)`[2],2)
  
  wald.text = ""
  
  if(!is.null(wald)){
    est = signif(wald[1],2)
    stder = signif(wald[2],2)
    t = signif(wald[3],2)
    wptext = ""
    if(!is.na(wald[4])){
      wptext = paste(", Wald p =",signif(wald[4],2))
    }
    wald.text = paste("beta = ",est,", std.err = ",stder, ", Wald t = ",t,wptext,';')
  }
  
  begin = 'There was no significant'
  if(p <0.1){
    begin = "There was a marginal"
  }
  if(p < 0.05){
    begin = 'There was a significant'  
  }
  
  
  return(paste(begin,ef,"(",wald.text,"log likelihood difference =",
               loglikDiff,", df = ",df,", Chi Squared =", chi,", p = ",p,")."))
}

```

# Load data

```{r}
variants = read.csv('../data/processedData/variants_summary.csv', stringsAsFactors = F)
```

There is only 1 variant for `white'.  Therefore, we remove it from this statistical analysis.

```{r}
variants = variants[variants$colourName!='white',]
```


Transform some variables.

```{r}

variants$Teach = variants$Teach >0

variants$check.any = variants$check>0

variants$freq_week_1.logcenter = log(variants$freq_week_1 + 1)
variants$freq_week_1.logcenter = 
  variants$freq_week_1.logcenter - mean(variants$freq_week_1.logcenter)

# cut TryMarking into two categories
variants$TryMarked.cat = cut(variants$TryMarked, 
                             c(-Inf,3,Inf), 
                             labels = c("Low",'High')) 

variants$averageLength_week_1.logcenter = log(variants$averageLength_week_1)
variants$averageLength_week_1.logcenter = 
  variants$averageLength_week_1.logcenter -
    mean(variants$averageLength_week_1.logcenter)

```

# ANOVA

Perform a straightforward ANOVA analysis:

```{r}
m1 = aov(freq_week_4_withinColour+1 ~ 
      indexical + 
      Teach * TryMarked*check.any+
      freq_week_1.logcenter +
      averageLength_week_1.logcenter+
      inventedBy + colourName,
    data=variants)
summary(m1)
```




# LMER models
```{r modelBits}
m0 = lmer(freq_week_4_withinColour ~ 
            1
          + (1 | colourName), 
          data=variants)
m1 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            + (1 | colourName), 
          data=variants)
m2 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach) 
          + (1 | colourName), 
          data=variants)

m3 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach) + (TryMarked.cat) 
          + (1 | colourName), 
          data=variants)

m4 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            Teach * TryMarked.cat
          + (1 | colourName), 
          data=variants)

m5 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter
          + (1 | colourName), 
          data=variants)

m6 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter
          + (1 | colourName), 
          data=variants)

m7 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any 
          + (1 | colourName), 
          data=variants)

m8 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any + inventedBy
          + (1 | colourName), 
          data=variants)

```

# Results

```{r}
anova(m0,m1,m2,m3,m4,m5, m6,m7,m8)
```

```{r}
summary(m7)
```


## Random slopes

For each of the predictors, we see if random slopes help improve the model.  Random slopes allow the strenght of the effect of a factor to be different for each colour concept.

```{r}

m7R = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any 
          + (1 | colourName), 
          data=variants)
          
m7R.indexical = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any 
          + (1 + indexical | colourName), 
          data=variants)

anova(m7R,m7R.indexical)

m7R.Teach = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any 
          + (1 + Teach| colourName), 
          data=variants)
anova(m7R,m7R.Teach)

m7R.TryMark = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any 
          + (1 + TryMarked.cat| colourName), 
          data=variants)
anova(m7R,m7R.TryMark)

m7R.Freq = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any 
          + (1 + averageLength_week_1.logcenter| colourName), 
          data=variants)
anova(m7R,m7R.Freq)

m7R.Length = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any 
          + (1 + averageLength_week_1.logcenter| colourName), 
          data=variants)
anova(m7R,m7R.Length)

m7R.Check = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any 
          + (1 + check.any| colourName), 
          data=variants)
anova(m7R,m7R.Check)
```

We see that only indexicality improves the model as a random slope.  This means that for some colours, indexicality matters more than for others.  We can plot the random slopes below:

```{r}
dotplot(ranef(m7R.indexical))
```

We see that indexicality is a better predictor of selection for black and red, but is a very weak predictor for pink and green.


# Summary

Choose a final model for the beta values:

```{r}
finalModel = m7R.indexical
```

`r getMEText(anova(m2,m3), "main effect of try marking",summary(finalModel)$coef['TryMarked.catHigh',])`

`r getMEText(anova(m3,m4), "interaction between try marking and teaching",summary(finalModel)$coef['TeachTRUE:TryMarked.catHigh',])`

`r getMEText(anova(m4,m5), "main effect of frequency in week 1",summary(finalModel)$coef['freq_week_1.logcenter',])`

`r getMEText(anova(m0,m1), "main effect of indexicality",summary(finalModel)$coef['indexicalYes',])`

`r getMEText(anova(m5,m6), "main effect of sign length",summary(finalModel)$coef['averageLength_week_1.logcenter',])`

`r getMEText(anova(m6,m7), "main effect of checking",summary(finalModel)$coef['check.anyTRUE',])`

`r getMEText(anova(m7,m8), "main effect of first user (inventedBy)")`

# Graphs

Plot the interaction between teaching and try marking.

```{r}

sumStats2 = summarySE(variants, measurevar="freq_week_4_withinColour",
                      groupvars=c("TryMarked.cat","Teach"))

dodge <- position_dodge(width=0.5) 

main.plot <- ggplot(sumStats2,
	aes(x = TryMarked.cat, y = freq_week_4_withinColour, colour=Teach)) +
  geom_point(position=dodge, size=4) + geom_line(aes(group=Teach),position=dodge) +
  geom_errorbar(aes(ymax=freq_week_4_withinColour+se, ymin=freq_week_4_withinColour-se), width=0.25,position=dodge) +
  xlab("Try Marked Frequency") +
  ylab("Proportional frequency in week 3") +
  coord_cartesian(ylim=c(0,0.6)) + 
  scale_color_discrete(breaks=c(FALSE,TRUE),
                       labels=c("No","Yes"),
                       name="Teach") +
  theme(text=element_text(size=18))
main.plot
```

```{r echo=F}
pdf("../results/descriptive/graphs/Interaction_Teach_TryMark.pdf", height=5, width=6)
main.plot
dev.off()
```

Plot the interaction between teaching and checking.

```{r}
sumStats2 = summarySE(variants, measurevar="freq_week_4_withinColour",
                      groupvars=c("Teach","check.any"))

dodge <- position_dodge(width=0.5) 

main.plot2 <- ggplot(sumStats2,
	aes(x = check.any, y = freq_week_4_withinColour, colour=Teach)) +
  geom_point(position=dodge, size=4) + geom_line(aes(group=Teach),position=dodge) +
  geom_errorbar(aes(ymax=freq_week_4_withinColour+se, ymin=freq_week_4_withinColour-se), width=0.25,position=dodge) +
  xlab("Check") +
  ylab("Proportional frequency in week 3") +
  coord_cartesian(ylim=c(0,0.25)) + 
  scale_color_discrete(breaks=c(FALSE,TRUE),
                       labels=c("No","Yes"),
                       name="Teach") +
  theme(text=element_text(size=18))
main.plot2
```


```{r echo=F}
pdf("../results/descriptive/graphs/Interaction_Teach_Check.pdf", height=5, width=6)
main.plot2
dev.off()
```