---
title: "Colour experiment"
output: pdf_document
---

```{r}

library(ggplot2)
library(lme4)
```

```{r echo=F}

setwd("~/Documents/MPI/KangSukColours/ColourExperiment/analysis/")

getMEText = function(r,ef, wald=NULL){
  
  AIC = r[2,]$AIC
  loglikDiff = signif(diff(r$logLik),2)
  chi = round(r$Chisq[2],2)
  df = r$`Chi Df`[2]
  p = signif(r$`Pr(>Chisq)`[2],2)
  
  wald.text = ""
  
  if(!is.null(wald)){
    est = signif(wald[1],2)
    stder = signif(wald[2],2)
    t = signif(wald[3],2)
    wptext = ""
    if(!is.na(wald[4])){
      wptext = paste(", Wald p =",signif(wald[4],2))
    }
    wald.text = paste("beta = ",est,", std.err = ",stder, ", Wald t = ",t,wptext,';')
  }
  
  begin = 'There was no significant'
  if(p <0.1){
    begin = "There was a marginal"
  }
  if(p < 0.05){
    begin = 'There was a significant'  
  }
  
  
  return(paste(begin,ef,"(",wald.text,"log likelihood difference =",
               loglikDiff,", df = ",df,", Chi Squared =", chi,", p = ",p,")."))
}

```

# Load data

```{r}
variants = read.csv('../data/processedData/variants_summary.csv', stringsAsFactors = F)

```

Transform some variables.

```{r}

variants$Teach = variants$Teach >0

variants$check.any = variants$check>0

variants$freq_week_1.logcenter = log(variants$freq_week_1 + 1)
variants$freq_week_1.logcenter = 
  variants$freq_week_1.logcenter - mean(variants$freq_week_1.logcenter)

# cut TryMarking into two categories
variants$TryMarked.cat = cut(variants$TryMarked, 
                             c(-Inf,3,Inf), 
                             labels = c("Low",'High')) 

variants$averageLength_week_1.logcenter = log(variants$averageLength_week_1)
variants$averageLength_week_1.logcenter = 
  variants$averageLength_week_1.logcenter -
    mean(variants$averageLength_week_1.logcenter)

```


# LMER models
```{r modelBits}
m0 = lmer(freq_week_4_withinColour ~ 
            1
          + (1 | colourName), 
          data=variants)
m1 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            + (1 | colourName), 
          data=variants)
m2 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach) 
          + (1 | colourName), 
          data=variants)

m3 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach) + (TryMarked.cat) 
          + (1 | colourName), 
          data=variants)

m4 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            Teach * TryMarked.cat
          + (1 | colourName), 
          data=variants)

m5 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter
          + (1 | colourName), 
          data=variants)

m6 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter
          + (1 | colourName), 
          data=variants)

m7 = lmer(freq_week_4_withinColour ~ 
            (indexical) + 
            (Teach * TryMarked.cat) +
            freq_week_1.logcenter +
            averageLength_week_1.logcenter+
            check.any 
          + (1 | colourName), 
          data=variants)

```

# Results

```{r}
anova(m0,m1,m2,m3,m4,m5, m6,m7)
```

```{r}
summary(m7)
```

## Summary

`r #getMEText(anova(m2,m3), "main effect of try marking",summary(m7)$coef['TryMarked.cat',])`

`r getMEText(anova(m4,m5), "main effect of frequency in week 1",summary(m7)$coef['freq_week_1.logcenter',])`

`r getMEText(anova(m0,m1), "main effect of indexicality",summary(m7)$coef['indexicalYes',])`

`r getMEText(anova(m5,m6), "main effect of sign length",summary(m7)$coef['averageLength_week_1.logcenter',])`


```{r}
#sjp.lmer(m5,'fe')
```

